Staged Abstract Interpreters

### 2019-09-07 paper/10_intro.tex
space

Roughly at the same time when abstract interpretation was proposed in the
1970s, \citet{futamura1971partial} observed a close
connection between interpreters and compilers through a hierarchy of
specializations, since then known as Futamura Projections. The first Futamura
projection describes that specializing an interpreter to the input program
removes interpretative overhead and yields an equivalent executable
that runs faster. 
In recent years, the idea of Futamura projection has regained more attentions
and successfully applied to many real-world problems, examples including
query compilers \cite{DBLP:conf/sigmod/TahboubER18}, PyPy \cite{Bolz:2009:TMP:1565824.1565827},
Truffle \cite{Marr:2015:TVP:2814270.2814275}, etc.
The challenge here: can we apply the same idea to a general abstract interpreter
to make it run faster? The solution is not obvious, as we need to specialize
a nonstandard semantics that extensively uses nondeterminism, operates on
abstract domains, and computes fixed-points iteratively over a cache.

>>>>

Roughly at the same time when abstract interpretation was proposed in the
1970s, \citet{futamura1971partial} observed a close
connection between interpreters and compilers through a hierarchy of
specializations, since then known as Futamura Projections. The first Futamura
projection describes that specializing an interpreter to the input program
removes interpretative overhead and yields an equivalent executable
that runs faster. 
In recent years, the idea of Futamura projections has regained more attentions
and successfully applied to many real-world problems, examples including
query compilers \cite{DBLP:conf/sigmod/TahboubER18}, PyPy \cite{Bolz:2009:TMP:1565824.1565827},
Truffle \cite{Marr:2015:TVP:2814270.2814275}, etc.
The challenge here: can we apply the same idea to a general abstract interpreter
to make it run faster? The solution is not obvious, as we need to specialize
a nonstandard semantics that extensively uses nondeterminism, operates on
abstract domains, and computes fixed-points iteratively over a cache.

### 2019-09-07 paper/20_prelim.tex
misc
space

The Scala compiler translates the above @for@-comprehension expression into
an equivalent one using @flatMap@ and @map@ ~\cite{scala_spec}. The last binding
in @for@-comprehensions is translated into a @map@, where the argument of
@yield@ becomes the body expression inside of that @map@ operation. The 
bindings before the last one are all translated into calls of @flatMap@:
\begin{lstlisting}
  xs.flatMap { case x => ys.map { case y => (x, y) } }
\end{lstlisting}

@@

Given the expression @e@, initial environment $\rho_0$ and store $\sigma_0$,
the top-level @run@ method invokes the fixed-point operator directly,
\begin{lstlisting}
  def run(e: Expr): Result = fix(eval)(e)(ρ$_0$)(σ$_0$)
\end{lstlisting}

>>>>

The Scala compiler translates the above @for@-comprehension expression into
an equivalent one using @flatMap@ and @map@. The last binding
in @for@-comprehensions is translated into a @map@, where the argument of
@yield@ becomes the body expression inside of that @map@ operation. The 
bindings before the last one are all translated into calls of @flatMap@:
\begin{lstlisting}
  xs.flatMap { case x => ys.map { case y => (x, y) } }
\end{lstlisting}

@@

Given the expression @e@, initial environment $\rho_0$ and store $\sigma_0$,
the top-level @run@ method invokes the fixed-point operator directly.
\begin{lstlisting}
  def run(e: Expr): Result = fix(eval)(e)(ρ$_0$)(σ$_0$)
\end{lstlisting}
